### 2013-10-22 article.tex
Made entire solution less drafty.

% INTRODUCTION
\section*{Introduction} % The \section*{} command stops section numbering
\addcontentsline{toc}{section}{\hspace*{-\tocsep}Introduction} % Adds this section to the table of contents with negative horizontal space equal to the indent for the numbered sections
\label{sec:introduction}
% Background and Challenge
%	Vilket sammanhang utgör studiens bakgrund?
%	Vilka utmaningar omfattar studien?
When developing \textit{GPU}-kernels, relatively small modifications in code may induce large deviations in performance due to massively parallelized instruction sets and architectural differences in-
between on-chip hardware. The increased utilization of complex \textit{GPGPU}-kernels has brought forth the need of more extensive debugging- and profiling-possibilities involving access of data that may be hard to retrieve from hardware. Therefore, it may be desirable for the developer to be able to view the data he or she is modifying on the graphics card - a possibility often limited in terms of \textit{GPGPU}-technologies, possibly due to architectural differences in-between chip manufacturers.\\
% Approach and Result
%	Vilken ansats tillämpas i studien?
A preferred solution to this problem has been to emulate such \textit{GPU}-kernels on the \textit{CPU}, such as described by Kerr et al.\cite[p.~416-419]{Hwu:2011:GCG:2103614} concerning the implementation of the \textit{GPU~Ocelot} compilation framework\footnote{\label{ftn:ocelot}\textit{GPU}~Ocelot: \url{http://b.gatech.edu/1a489L1}. Retrieved 22-10-2013.}, often in exchange for substantial performance-losses. Other resons to emulate \textit{GPU}-kernels may concern pre-silicon development - that is, development for hardware not yet existant, when hardware is busy, or otherwize unavailable\footnote{\label{ftn:warpguide}Microsoft's reference on \textit{WARP}: \url{http://bit.ly/19hiwrZ}. Retrieved: 21-10-2013.}. This study comprises an investigation into the performance of software emulation of hardware accelerated \textit{GPGPU}-kernels, by the means of analyzing several software rasterizers.\\
\\
Furthermore, this material concerns inquiry into the \textit{DirectCompute}-framework on the \textit{Windows}-platform, analyzing the performance of a \textit{GPGPU}-kernel on-chip, using the \textit{DirectX} standard software rasterizer, and utilizing the \textit{DirectX~11.1}-addition Microsoft~\textit{WARP}\footnote{\textit{Windows~Advanced~Rasterization~Platform}.}-technology - which promises high-speed software emulation\footref{ftn:warpguide}. \\
\\
%...in exchange for some computational precision. Not Supported by cited source. From where did I get that idea?
%	Vilka resultat gav studien?
This article concludes that the performance losses inflicted by such emulation may be reduced enough for that emulation to be considered feasable for computational use in real-time, as originally proposed by Microsoft\footref{ftn:warpguide}. Thus, this study proposes use of Microsoft~\textit{WARP}-technology in industry rasterization if graphics~hardware is unavailable, not sufficient, or busy.\\
\\
% Outline and Conclusion
As such, the study concerns the fields of simulation, emulation and \textit{GPU}-technologies - respectively, with the purpose of facilitating debugging and profiling of \textit{GPU}-kernels, whilst maintaining acceptable performance. \\
The remainder of this document presents the method and process to acquire the data used,  the technologies using which it has been acquired, the results in and of their own, the conclusions based off the results and finally; the author's personal reflections surrounding future~work in the area along with propositions of further study.
%	Vilket upplägg har resterande delar av matrialet?

@@

\subsection{Subjects~of~Study}
\label{sec:contribution:Subjectofstudy}
% Describe these modes and what they entail:
The \textit{GPU}-kernels described in Section~\ref{sec:contribution:method} are comprised of \textit{HLSL}-syntax and are compiled \& executed using Microsoft~\textit{DirectCompute}. These kernels are run using three types of acceleration~technologies with varying strategies of \textit{GPU}-kernel emulation. These are comprised of the following:
\paragraph{Hardware-Acceleration~(GPU)}
	The execution of a \textit{DirectCompute}-kernel on a graphics card. This is the common case, and involves no emulation of the kernel. Thus, this case will act as a reference for the emulated Subjects.\\
	During hardware~acceleration on the \textit{GPU}, one may expect high performance.
\paragraph{Software~Rasterization~(CPU)}
	The emulated execution of a \textit{DirectCompute}-kernel using the \textit{DirectX}~Reference~Device~Driver.\\
The Reference~Driver was developed for the purpose of testing and debugging, and is - although it does support some \textit{CPU}-optimizations - not intended to be used in retail applications\footnote{\label{ftn:drivertypes}Microsoft's reference on \textit{DirectX}~Driver~Types: \url{http://bit.ly/1b63NiX}. Retrieved: 21-10-2013.}.\\
	As the Reference~Driver is designed for the purpose of accuracy, rather than speed, one may expect poor performance.
\paragraph{Windows~Advanced~Rasterization~Platform~(CPU)}
	The emulated execution of a \textit{DirectCompute}-kernel using a special software rasterizer devised by Microsoft in their latest revision of the \textit{DirectX}-framework.\\
	The driver is based off the \textit{DirectX}~Reference~Driver and uses thread pooling to distribute tasks efficiently on the \textit{CPU}, along with grouping execution in batches for optimum performance. Microsoft describes \textit{WARP} as a high-performance software~rasterizer, and recommends using the driver for retail applications, such as casual~games\footref{ftn:warpguide}.\\
As there are, as of yet, few studies performed on Microsoft~\textit{WARP} expectations are unclear, but the driver is expected to perform better than standard software rasterization.\\
% May one encounter computational precision issues with WARP?

@@

% CONCLUSION
\section{Conclusion}
\label{sec:conclusion}
Based off the results presented in Section~\ref{sec:contribution:results}, it would appear as if Microsoft~\textit{WARP} is an impressive driver in comparison to the \textit{DirectX}~Reference~Device, from which it has been derived. Using \textit{WARP} to emulate the kernels presented in this study has magnitudes greater performance than if one were to apply the \textit{DirectX}~Reference~Device in the same manner. If one were to compare the execution of \textit{WARP}~\&~the~Reference~Device side-by-side, and assume the same area of application, \textit{WARP} is superior for the purposes presented in this document.\\
However, keeping in mind the major performance improvements offered by Microsoft~\textit{WARP}, it is important to consider that the two may be appropriate for different purposes. The \textit{DirectX}~Reference~Device is primarily proposed by Microsoft as a debugging/pre-silicon-development tool, whereas \textit{WARP} is intended for use in a broader sense\footref{ftn:warpguide} - such as to render graphics for casual games - in addition to debugging and error-profiling purposes.\\
\\
In regards to the decreased performance of \textit{WARP}-accelerated kernels when utilizing shared memory; it is probable that... \\
\\
In conclusion; this study proposes, pursuant to the established performance of Microsoft~\textit{WARP}, that \textit{WARP}-technologies are feasable for extended use in applications - for purposes other than debugging and profiling.

@@

% CONCLUSION - FUTURE WORK
\subsection{Future~Work}
\label{sec:conclusion:futurework}
For the sake of brevity, the author suggests complementary elaboration into double precision calculations in \textit{DirectCompute}-kernels, with the intent of examining whether or not the Subjects detailed in Section~\ref{sec:contribution:Subjectofstudy} may have differentiating effects on computational precision.

@@

% ACKNOWLEDGMENTS
\section*{Acknowledgments} % The \section*{} command stops section numbering
\addcontentsline{toc}{section}{\hspace*{-\tocsep}Acknowledgments} % Adds this section to the table of contents with negative horizontal space equal to the indent for the numbered sections
\label{sec:acknowledgments}
%I wish to express my gratitude to my fellow students, colleagues and close friends Bob, Benny and no-one for their work, co-operation and positive attitude throughout the execution of this experiment.

>>>>

% INTRODUCTION
\section*{Introduction} % The \section*{} command stops section numbering
\addcontentsline{toc}{section}{\hspace*{-\tocsep}Introduction} % Adds this section to the table of contents with negative horizontal space equal to the indent for the numbered sections
\label{sec:introduction}
% Background and Challenge
%	Vilket sammanhang utgör studiens bakgrund?
%	Vilka utmaningar omfattar studien?
When developing \textit{GPU}-kernels, relatively small modifications in code may induce large deviations in performance due to massively parallelized instruction sets and architectural differences in-
between on-chip hardware\footnote{See Performance~Considerations by Kirk~\&~Hwu\cite[ch.~6]{Kirk:2010:PMP:1841511} for an analyis on the volatility of \textit{GPGPU}-performance.}. The increased utilization of complex \textit{GPGPU}-kernels has brought forth the need of more extensive debugging- and profiling-possibilities involving access of data that may be hard to retrieve from hardware. Therefore, it may be desirable for the developer to be able to view the data he or she is modifying on the graphics card - a possibility often limited in terms of \textit{GPGPU}-technologies, possibly due to architectural differences in-between chip manufacturers.\\
% Approach and Result
%	Vilken ansats tillämpas i studien?
A preferred solution to this problem has been to emulate such \textit{GPU}-kernels on the \textit{CPU}\footnote{\label{ftn:drivertypes}Microsoft's reference on \textit{DirectX}~Driver~Types: \url{http://bit.ly/1b63NiX}. Retrieved: 21-10-2013.}, such as described by Kerr et al.\cite[p.~416-419]{Hwu:2011:GCG:2103614} concerning the implementation of the \textit{GPU~Ocelot} compilation framework\footnote{\label{ftn:ocelot}\textit{GPU}~Ocelot: \url{http://b.gatech.edu/1a489L1}. Retrieved 22-10-2013.}, often in exchange for substantial performance-losses. Other resons to emulate \textit{GPU}-kernels may concern pre-silicon development - that is, development for hardware not yet existant, when hardware is busy, or otherwize unavailable\footnote{\label{ftn:warpguide}Microsoft's reference on \textit{WARP}: \url{http://bit.ly/19hiwrZ}. Retrieved: 21-10-2013.}. This study comprises an investigation into the performance of software emulation of hardware accelerated \textit{GPGPU}-kernels, by the means of analyzing several software rasterizers.\\
\\
Furthermore, this material concerns inquiry into the \textit{DirectCompute}-framework on the \textit{Windows}-platform, analyzing the performance of a \textit{GPGPU}-kernel on-chip, using the \textit{DirectX} standard software rasterizer, and utilizing the \textit{DirectX~11.1}-addition Microsoft~\textit{WARP}\footnote{\textit{Windows~Advanced~Rasterization~Platform}.}-technology - which promises high-speed software emulation\footref{ftn:warpguide}. \\
\\
%...in exchange for some computational precision. Not Supported by cited source. From where did I get that idea?
%	Vilka resultat gav studien?
This article concludes that the performance losses inflicted by such emulation may be reduced enough for that emulation to be considered feasable for computational use in real-time, as originally proposed by Microsoft\footref{ftn:warpguide}. Thus, this study proposes use of Microsoft~\textit{WARP}-technology in industry rasterization if graphics~hardware is unavailable, not sufficient, or busy.\\
\\
% Outline and Conclusion
As such, the study concerns the fields of simulation, emulation and \textit{GPU}-technologies - respectively, with the purpose of facilitating debugging and profiling of \textit{GPU}-kernels, whilst maintaining acceptable performance. \\
The remainder of this document presents the method and process to acquire the data used,  the technologies using which it has been acquired, the results in and of their own, the conclusions based off the results and finally; the author's personal reflections surrounding future~work in the area along with propositions of further study.
%	Vilket upplägg har resterande delar av matrialet?

@@

\subsection{Subjects~of~Study}
\label{sec:contribution:Subjectofstudy}
% Describe these modes and what they entail:
The \textit{GPU}-kernels described in Section~\ref{sec:contribution:method} are comprised of \textit{HLSL}-syntax and are compiled \& executed using Microsoft~\textit{DirectCompute}. These kernels are run using three types of acceleration~technologies with varying strategies of \textit{GPU}-kernel emulation. These are comprised of the following:
\paragraph{Hardware-Acceleration~(GPU)}
	The execution of a \textit{DirectCompute}-kernel on a graphics card. This is the common case, and involves no emulation of the kernel. Thus, this case will act as a reference for the emulated Subjects.\\
	During hardware~acceleration on the \textit{GPU}, one may expect high performance.
\paragraph{Software~Rasterization~(CPU)}
	The emulated execution of a \textit{DirectCompute}-kernel using the \textit{DirectX}~Reference~Device~Driver.\\
The Reference~Driver was developed for the purpose of testing and debugging, and is - although it does support some \textit{CPU}-optimizations - not intended to be used in retail applications\footref{ftn:drivertypes}.\\
	As the Reference~Driver is designed for the purpose of accuracy, rather than speed, one may expect poor performance.
\paragraph{Windows~Advanced~Rasterization~Platform~(CPU)}
	The emulated execution of a \textit{DirectCompute}-kernel using a special software rasterizer devised by Microsoft in their latest revision of the \textit{DirectX}-framework.\\
	The driver is based off the \textit{DirectX}~Reference~Driver and uses thread pooling to distribute tasks efficiently on the \textit{CPU}, along with grouping execution in batches for optimum performance. Microsoft describes \textit{WARP} as a high-performance software~rasterizer, and recommends using the driver for retail applications, such as casual~games\footref{ftn:warpguide}.\\
As there are, as of yet, few studies performed on Microsoft~\textit{WARP} expectations are unclear, but the driver is expected to perform better than standard software rasterization.\\
% May one encounter computational precision issues with WARP?

@@

% CONCLUSION
\section{Conclusion}
\label{sec:conclusion}
Based off the results presented in Section~\ref{sec:contribution:results}, using \textit{WARP} to emulate the kernels presented in this study has magnitudes greater performance than if one were to apply the \textit{DirectX}~Reference~Device in the same manner. Hence, if one were to compare the execution of \textit{WARP}~\&~the~Reference~Device side-by-side, and assume the same area of application, \textit{WARP} is superior in terms of execution time - assuming the same preconditions as those presented in this material.\\
However, keeping in mind the major performance improvements offered by Microsoft~\textit{WARP}, it is important to consider that the two may be appropriate for different purposes. The \textit{DirectX}~Reference~Device is primarily proposed by Microsoft as a debugging/pre-silicon-development tool, whereas \textit{WARP} is intended for use in a broader sense\footref{ftn:warpguide} - such as to render graphics for casual games - in addition to debugging and error-profiling purposes.\\
This calls for further inquiry into what the flaws of using Microsoft~\textit{WARP} may be.\\
\\
In conclusion; this study proposes, pursuant to the established performance of Microsoft~\textit{WARP}, that \textit{WARP}-technologies are feasable for extended use in applications - for purposes other than debugging and profiling.

%In regards to the decreased performance of \textit{WARP}-accelerated kernels when utilizing shared memory; it is probable that...

@@

% CONCLUSION - FUTURE WORK
\subsection{Future~Work}
\label{sec:conclusion:futurework}
For the sake of brevity, the author suggests complementary elaboration into double precision calculations in \textit{DirectCompute}-kernels, with the intent of examining whether or not the Subjects detailed in Section~\ref{sec:contribution:Subjectofstudy} may have differentiating effects on computational precision.\\
\\
Additionally, the use of more complex kernels  in coagency with the \textit{WARP}-driver should be examined in order to study the effects of more demanding emulation on the \textit{CPU}.

@@

% ACKNOWLEDGMENTS
%\section*{Acknowledgments} % The \section*{} command stops section numbering
%\addcontentsline{toc}{section}{\hspace*{-\tocsep}Acknowledgments} % Adds this section to the table of contents with negative horizontal space equal to the indent for the numbered sections
%\label{sec:acknowledgments}
%I wish to express my gratitude to my fellow students, colleagues and close friends Bob, Benny and no-one for their work, co-operation and positive attitude throughout the execution of this experiment.
